---
title: "p8105_hw2_xh2395"
author: "Xin  He"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readxl)
```


## Problem 1
**Read and clean the Mr. Trash Wheel sheet:**
```{r}
Mr_Trash_data = 
  read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "Mr. Trash Wheel", range = cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  mutate(sports_balls = as.integer(round(sports_balls)))
```

**Read and clean precipitation data for 2018**
```{r}
preci2018_data = 
  read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = 2018)
```

**Read and clean precipitation data for 2017**
```{r}
preci2017_data = 
  read_excel(path = "./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = "2017 Precipitation", range = "A2:B14") %>% 
  janitor::clean_names() %>%
  drop_na(total) %>% 
  mutate(year = 2017)
```

**Combine precipitation datasets**
```{r}
com_data = full_join(preci2017_data, preci2018_data, by = NULL) %>% 
  mutate(month = month.name[month])
```

In dataset "Mr_Trash_data", the number of observations is `r nrow(Mr_Trash_data)`, and the number of variables is `r ncol(Mr_Trash_data)`. Key variables include "dumpster" and "sports_balls". 

In dataset "preci2018_data", the number of observations is `r nrow(preci2018_data)`, and the number of variables is `r ncol(preci2018_data)`. Key variables include "year", "month" and "total".

In dataset "preci2017_data", the number of observations is `r nrow(preci2017_data)`, and the number of variables is `r ncol(preci2017_data)`. Key variables include "year", "month" and "total".

In dataset "com_data", the number of observations is `r nrow(com_data)`, and the number of variables is `r ncol(com_data)`. Key variables include "year", "month" and "total".

The total precipitation in 2018 is `r sum(0.96+5.30+2.18+3.20+9.27+0.20+2.39)` inches.

The median number of sports balls in a dumpster in 2017 is `r median(Mr_Trash_data$ sports_balls)`.


## Problem 2
**Clean the data in pols-month.csv**
```{r}
pols_data = 
  read_csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>% 
  mutate(month = month.name[month]) %>% 
  mutate(president = ifelse(prez_gop %in% 0, "dem",
     ifelse(prez_gop %in% 1, "gop", "gop"))) %>% 
  select(-prez_gop, -prez_dem, -day)
```

**Clean the data in snp.csv**
```{r}

```














